Sparse Word Embedding for Multi-domain Machine Translation

### 2019-04-08 2019/emnlp/sparse.tex
wmt,emnlp

\documentclass[11pt,a4paper]{article}
\usepackage[hyperref]{naaclhlt2019}
\usepackage{times}
\usepackage{latexsym}
\usepackage{graphicx}
\usepackage{multirow}
\usepackage{url}
\usepackage{float} 
%\aclfinalcopy % Uncomment this line for the final submission
%\def\aclpaperid{***} %  Enter the acl Paper ID here

@@

\newcommand\BibTeX{B{\sc ib}\TeX}

@@

\begin{document}
\maketitle
\begin{abstract}
In this paper, we introduce an application of an old technique \cite{P07-1033} in Multi-Domain Neural Machine Translation.  By defining general information encoding region and domain-specific encoding region in word embedding vector, we are able to  seperate domain-related information flows in forward and backward propagation during training
mitigate the catastrophic interference happens between the inference of different domains in Neural Machine Translation system, i.e to improve the performance of the system for every domain in which it was trained. 
\end{abstract}

@@

\section{Related Work}
Multi-domain Machine Translation has largely interested NLP community by its promising applications in industry and its relation to fundamental problems in theory. Researchers have proposed a large range of techniques from Data centric methods to Model centric methods \cite{C18-1111}; \cite{P17-2061}. Data centric methods have goal to collect related-domain data from existing in-domain data by using different techniques such as scoring relatedness by Language Model (\cite{P10-2041}; \cite{D11-1033}; \cite{P13-2119} or by metric on the space of sentence embedding \cite{P17-2089} or generating pseudo parallel data \cite{P03-1010}; \cite{C16-1295}; \cite{D14-1023}. On the other hand, Model centric approaches focus on NMT models that are specialized for domain adaptation. The novelty can be either trainining objective \cite{Luong2015SNMT}; \cite{P16-1009}; \cite{D17-1155}; \cite{W17-3205}; \cite{D17-1156}; \cite{C18-1269} or architecture \cite{R17-1049}; \cite{gulcehre2016monolingual}; \cite{W17-4712}, \cite{Biao2017CARENMT}; \cite{N18-2080}; \cite{W18-6313}; \cite{C16-1170}; \cite{P18-2050} or the decoding algorithm \cite{gulcehre2016monolingual}; \cite{I17-2004}. Beside these works, we could also consider "out-of-domain" works such as sequence labeling tasks \cite{P07-1033}; learning multiple visual domains \cite{NIPS2017_6654}. The problem is recently investigated by several interesting works presented in EMNLP 2018 such as \cite{D18-1039}; \cite{D18-1041}. \cite{D18-1041} create in the encoder Domain-specific gate $g^r_i$and Domain-shared gate $g^s_i$ which are generated from domain-specific and domain-shared semantic representations of source sentence $E_r(x)$ and $E_s(x)$ respectively and which select information from units of hidden states $h_i$ by elementwise product $h^r_i = g^r_i \odot h_i$; $h^s_i = g^s_i \odot h_i$. $h^r_i$ and $h^s_i$ will be fed to Domain-specific and Domain-shared attentional mechanisms respectively. While introducing new features in the architecture, \cite{D18-1041} introduces also new objective which is sum of word-level weighted MT objective and objectives of Domain-classifier in source side and target side and Adversarial Domain-classifier in source side. The author have very good approach to well seperate Domain-shared information flow and Domain-specify information flow and determine their contributions to the inference that mitigates the catastrophic interference happens in the network during forward step.

\section{The Model}
There is not any changes in the architecture of Neural Machine Translation network except the construction of word embedding. 
\begin{figure}[h]
\center
    \includegraphics[width=0.8\linewidth]{Sparse1}
    \caption{Illustration of the model.} 
    \label{network}
\end{figure}
\section{Experiments}

\subsection{Corpora}
To evaluate the performance of proposed architecture, we use 4 corpora for training: European Medicines Agency; European Central Bank, News-commentary and European Parliament \cite{Tiedemann2009RANLP5} corresponding to 4 domains: medical, banking, news and administrative  respectively and one non-domain corpora Common-crawl  which was introduced in Shared Task: Machine Translation of News the Third Conference On Machine Translation 2018(WMT2018). We use Khresmoi-dev, a test set used in the Biomedical Translation Task (WMT2018), newstest2008, and test2006 as validation set for domains: medical, news administrative; except for banking domain, we use a subset of size 2000 extracted from original corpora and excluded from training set. To test the model, we use Khresmoi-test; newstest2009 and test2007 for domains: medical, news administrative; except for banking domain, we use a subset of size 2000 extracted from original corpora and excluded from training set. The information of used corpora is presented in table \ref{table:Corpora}
\begin{table}[H]
\resizebox{\columnwidth}{!}{
\begin{tabular}{ |l|l|l|l|l|l| }
\hline
Task & Corpora & Train & Dev & Test & Vocabulary \\ \hline
\multirow{5}{*}{$English \rightarrow French$} & European Medicines Agency & 1092568 & 500 & 1000 & \\
 & European Central Bank & 195960 & 2000 & 2000 &\\
 & News-commentary & 258432 & 2051 & 3027 & \\
 & European Parliament & 2007723 & 2000 & 2000 & \\
 & Common-crawl & 3244152 &  &  & \\ \hline
\multirow{5}{*}{$English \rightarrow German$} & European Medicines Agency & 1108752 & 500 & 1000 & \\
 & European Central Bank & 113174 & 2000 & 2000 &\\
 & News-commentary & 270769 & 2051 & 2525 & \\
 & European Parliament & 1920209 & 2000 & 2000 & \\ 
 & Common-crawl & 2399123 &  &  & \\ \hline
\hline
\end{tabular}
}
\caption{Corpora}
\label{table:Corpora}
\end{table}

@@

\subsection{Results}

\section{Conclusions}

@@

\bibliography{naaclhlt2019}
\bibliographystyle{acl_natbib}

>>>>

\documentclass[11pt,a4paper]{article}
\usepackage[hyperref]{naaclhlt2019}
\usepackage{times}
\usepackage{amsmath,amsthm,amsfonts,amssymb,amscd}
\usepackage{latexsym}
\usepackage{graphicx}
\usepackage{multirow}
\usepackage{url}
\usepackage{float} 
\usepackage{tabularx}
\newtheorem{lemma}{Lemma}[section]
\newtheorem{theorem}{Theorem}[section]
\newtheorem{proposition}{Proposition}[section]
\newtheorem{definition}{Definition}[section]
\DeclareMathOperator*{\argmax}{argmax}
%\aclfinalcopy % Uncomment this line for the final submission
%\def\aclpaperid{***} %  Enter the acl Paper ID here

@@

@@

\begin{document}
\maketitle
\begin{abstract}
In this paper, we introduce an application of an old technique \cite{P07-1033} in Multi-Domain Neural Machine Translation.  By defining general information encoding region and domain-specific encoding region in word embedding vector, we are able to  seperate domain-related information flows in forward and backward propagation during training thus to
mitigate the catastrophic interference happens between the inference of different domains in Neural Machine Translation system, i.e to improve the performance of the system for every domain in which it was trained. 
\end{abstract}

@@

\section{Related Work}
Multi-domain Machine Translation has largely interested NLP community by its promising applications in industry and its relation to fundamental problems in theory. Researchers have proposed a large range of techniques from Data centric methods to Model centric methods \cite{C18-1111}; \cite{P17-2061}. Data centric methods have goal to collect related-domain data from existing in-domain data by using different techniques such as scoring relatedness by Language Model (\cite{P10-2041}; \cite{D11-1033}; \cite{P13-2119} or by metric on the space of sentence embedding \cite{P17-2089} or generating pseudo parallel data \cite{P03-1010}; \cite{C16-1295}; \cite{D14-1023}. On the other hand, Model centric approaches focus on NMT models that are specialized for domain adaptation. The novelty can be either trainining objective; for example \cite{Luong2015SNMT}; \cite{P16-1009}; \cite{D17-1155}; \cite{W17-3205}; \cite{D17-1156}; \cite{C18-1269} or architecture \cite{R17-1049}; \cite{gulcehre2016monolingual}; \cite{W17-4712}, \cite{Biao2017CARENMT}; \cite{N18-2080}; \cite{W18-6313}; \cite{C16-1170}; \cite{P18-2050} or the decoding algorithm \cite{gulcehre2016monolingual}; \cite{I17-2004}. Beside these works, we could also consider "out-of-domain" works such as sequence labeling tasks \cite{P07-1033}; learning multiple visual domains \cite{NIPS2017_6654}. The problem is recently investigated by several interesting works presented in EMNLP 2018 such as \cite{D18-1039}; \cite{D18-1041}. \cite{D18-1041} create in the encoder Domain-specific gate $g^r_i$and Domain-shared gate $g^s_i$ which are generated from domain-specific and domain-shared semantic representations of source sentence $E_r(x)$ and $E_s(x)$ respectively and which select information from units of hidden states $h_i$ by elementwise product $h^r_i = g^r_i \odot h_i$; $h^s_i = g^s_i \odot h_i$. $h^r_i$ and $h^s_i$ will be fed to Domain-specific and Domain-shared attentional mechanisms respectively. While introducing new features in the architecture, \cite{D18-1041} introduces also new objective which is sum of word-level weighted MT objective and objectives of Domain-classifier in source side and target side and Adversarial Domain-classifier in source side. The author have very good approach to well seperate Domain-shared information flow and Domain-specify information flow and determine their contributions to the inference that mitigates the catastrophic interference happens in the network during forward step.

\section{The Model}
\begin{definition}{Domain}
\label{def:1}
Domain is a hypothetical probability that generates a set of corpora. We will denote D(C) is domain of set of corpora C.
\end{definition}

In machine translation, the loss function for training is
\begin{equation}
\begin{split}
E_{(x,y) \sim p_{rel}}[-log(p_{\theta}(y|x))] &= \\
= E_{x \sim p_{rel}(x)}[KL(p_{rel}(y|x) & \mid p_{\theta}(y|x))] + \\
+ E_{x \sim p_{rel}} &[H(p_{rel}(y|x))]
\end{split}
\end{equation}.
Where $p_{\theta}(y|x)) = p(y|x,\theta)$ or $p_{\theta}(y|x)) = p(y,\theta | x)$
 In this quantity, domain is exactly $p_{rel}(x,y)$. However, in the inference or translation, the most important goal is to predict target sentence given source sentence, i.e, to well approximate the conditional probability $p_{rel}(y|x)$. However, this relation could be very different if consider 2 domains generating 2 corpora of different topic. Indeed, in financial document, to translate "security" into "titre financier" is more likely than to translate it into "securit\'e" but in IT document. In reality, to translate correctly with respect to the topic of sentence is very important. We formalize the multi-domain problem for machine translation in following equation
\begin{equation}
\begin{split}
E_{(x,y) \sim D(C_{i})}[-log(p_{\theta^*}(y|x))] & = \\
= \argmax_{\theta} E_{(x,y) \sim D(C_{i})} [-log & (p_{\theta}(y|x))] \; \forall i \in [1..d]
\end{split}
\end{equation}
A trivial solution for this equation is $\theta^* = (\theta^*_{0},...,\theta^*_{d})$ and $p_{(\theta_{0},...,\theta_d)}(y|x,i)$ is independent of $\theta_{j}$ $\forall j \neq i$, i.e, $\exists f:\Omega_i \times C(D_i) \rightarrow [0,1]: f_{\theta_i}(y,x) = p_{(\theta_{0},...,\theta_d)}(y|x,i)$ where $\Omega_i$ is set of all possible values of $\theta_i$. \\
The model above is not different than ensemble of $d$ models specialized for one domain. However, we could mitigate this seperation by using shared paramaters denoted $\theta_s$ then reformulate the model as following: \\
$\theta^* = (\theta^*_{0},...,\theta^*_{d}, \theta_s^*)$ and $p_{(\theta_{0},...,\theta_d, \theta_s)}(y|x,i)$ is independent of $\theta_{j}$ $\forall j \neq i$, i.e, $\exists f:\Omega_i \times \Omega_s \times C(D_i) \rightarrow [0,1]: f_{\theta_i, \theta_s}(y,x) = p_{(\theta_{0},...,\theta_d, \theta_s)}(y|x,i)$ where $\Omega_i$ is set of all possible values of $\theta_i$ and $\Omega_s$ is set of all possible values of $\theta_s$. \\
However, there is a problem that the couple $\theta_s^*$ and $\theta_i^*$ have to be optimal $\forall i \in [1..d]$.\\
Our first approach to multi-domain machine translation is to build such trivial solution in calibrating positions and sizes of domain paramters $\theta_i$ and shared parameters $\theta_s$ inside one architecture which is harly achieved. Despite that fact, we could consider $\theta_i$ is additional factor used for fine-tuning the model to the domain $D(C_{i})$, i.e, lowering the perplexity with respect to true distribution of domain $E_{(x,y) \sim D(C_{i})}[-log(p_{\theta}(y|x,i))]$.
\begin{figure}[h]
\center
    \includegraphics[width=0.8\linewidth]{Sparse1}
    \caption{Illustration of the model.} 
    \label{network}
\end{figure}
\section{Experiments}

\subsection{Corpora}
To evaluate the performance of proposed architecture, we use 4 corpora for training: European Medicines Agency; European Central Bank, News-commentary and European Parliament \cite{Tiedemann2009RANLP5} corresponding to 4 domains: medical, banking, news and administrative  respectively and one non-domain corpora Common-crawl  which was introduced in Shared Task: Machine Translation of News the Third Conference On Machine Translation 2018(WMT2018). We use Khresmoi-dev, a test set used in the Biomedical Translation Task (WMT2018), newstest2008, and test2006 as validation set for domains: medical, news administrative; except for banking domain, we use a subset of size 2000 extracted from original corpora and excluded from training set. To test the model, we use Khresmoi-test; newstest2009 and test2007 for domains: medical, news administrative; except for banking domain, we use a subset of size 2000 extracted from original corpora and excluded from training set. The information of used corpora is presented in table \ref{table:Corpora}
\begin{table}[H]
\resizebox{\columnwidth}{!}{
\begin{tabular}{ |l|l|l|l|l|l| }
\hline
Task & Corpora & Train & Dev & Test & Vocabulary \\ \hline
\multirow{5}{*}{$English \rightarrow French$} & European Medicines Agency & 1092568 & 500 & 1000 & \\
 & European Central Bank & 191960 & 2000 & 2000 &\\
 & News-commentary & 258432 & 2051 & 3027 & \\
 & European Parliament & 2007723 & 2000 & 2000 & \\
 & Common-crawl & 3244152 &  &  & \\ \hline
\multirow{5}{*}{$English \rightarrow German$} & European Medicines Agency & 1108752 & 500 & 1000 & \\
 & European Central Bank & 109174 & 2000 & 2000 &\\
 & News-commentary & 270769 & 2051 & 2525 & \\
 & European Parliament & 1920209 & 2000 & 2000 & \\ 
 & Common-crawl & 2399123 &  &  & \\ \hline
\hline
\end{tabular}
}
\caption{Corpora}
\label{table:Corpora}
\end{table}

@@

\subsection{Results}
\begin{figure}[h]
\caption{BLEU score in language pair English - French}
\begin{center}
 \resizebox{\columnwidth}{!}{\begin{tabularx}{\textwidth}{|| X | X | X | X | X | X | X | X ||} 
 \hline
 Models & EMEA test & EPPS test & ECB test & khresmoi & test 2007 & newstest 2009 & IWSLT 2010 \\ [0.5ex] 
 \hline\hline
 Mixed & & & & & & & \\
 \hline
 EMEA-finetuned & & & & & & & \\
 \hline
 EPPS-finetuned & & & & & & & \\
 \hline
 ECB-finetuned & & & & & & & \\
 \hline
 Sparse & & & & & & & \\
 \hline
 Domain Control & & & & & & & \\
 \hline
\end{tabularx}}
\end{center}
\label{tab:1}
\end{figure}

\begin{figure}[h]
\caption{BLEU score in language pair English - German}
\begin{center}
 \resizebox{\columnwidth}{!}{\begin{tabularx}{\textwidth}{|| X | X | X | X | X | X | X | X ||} 
 \hline
 Models & EMEA test & EPPS test & ECB test & khresmoi & test 2007 & newstest 2009 & IWSLT 2010 \\ [0.5ex] 
 \hline\hline
 Mixed & & & & & & & \\
 \hline
 EMEA-finetuned & & & & & & & \\
 \hline
 EPPS-finetuned & & & & & & & \\
 \hline
 ECB-finetuned & & & & & & & \\
 \hline
 Sparse & & & & & & & \\
 \hline
 Domain Control & & & & & & & \\
 \hline
\end{tabularx}}
\end{center}
\label{tab:2}
\end{figure}
\section{Conclusions}

@@

\bibliography{naaclhlt2019}
\bibliographystyle{plainnat}
